\vspace{-4em}Rethinking Model Architecture and Means of Delivery to get Video Interaction with Generative Neural Networks Online

### 2020-11-26 writing/paper/draft_02/draft_02.tex
Final tweaks before handin

Unfortunately the gallery in itself is a black-box to many, entry restricted by a fee or simply by the location of the gallery. Moving such interactive experiences online is a clear next step as once something is online it is immediately available to anyone with an internet connection, however this then poses numerous challenges. Even a modestly sized NN can contain many millions of parameters. Modern computers are able to perform millions of floating point operations per second (FLOPS) but advances in ML research have shown that \textit{deeper} models and more data produces better results\cite{2005.14165}; as such, without expensive hardware acceleration from discrete graphics processing units (GPU) or even more modern tensor processing units (TPU), an average modern computer cannot crunch the numbers needed to run a large, deep NN\@. Even if one \textit{could} perform the necessary FLOPS, storing this number of parameters can take up gigabytes of memory and so downloading the weights and biases of a pre-trained model can be a time consuming process detracting from the real-time focus.

@@

Getting NNs working online and on the clients machine is not a new thing and with modern CPUs it has certainly been shown possible to train and run a simple enough model in the browser\cite{karparthy_conjs_2016, brainjs, synaptic, mind, kerasjs} and many tools exist which aim to visualise the architecture of a model and show the training process\cite{broad_cnn_vis_2016, tensorflow_playground, 1704.01942}. Such tools are invaluable as learning resources for those dedicated to learning about machine learning. However they are often tailored to that use-case expecting the user to have higher motivation for understanding the process, more than the average online `passer-by'.

@@

\textit{Learning to See} is an adaptation of the Pix2Pix model\cite{1611.07004} which is a Conditional Generative Adversarial Network (cGAN)\cite{1701.00160}. Unlike a conventional GAN\cite{1406.2661}, a cGAN takes in two data samples as input, in the case of the Pix2Pix model it is a \textit{target} image and an \textit{input} image. Pix2Pix also learns a structured loss which means it is able to penalize any possible structure that differs between output and target\cite{1611.07004}. This model architecture is extremely flexible and has been shown to be suitable to many use cases as detailed in the original paper. However neither the original Pix2Pix authors nor \textit{Learning to See} had an incentive to create a small model; a good result is simply a sharp image with a predictable output from a given input. As stated by the Pix2Pix authors measuring the results of an image generating GAN is an open issue\cite{1606.03498, 1611.07004} and in an artistic context quantifying the outcome is largely pointless --- a good result is one which delivers the message of the artist.

It is worth noting that other generative models of similar architecture to Pix2Pix have been developed since, notably CycleGAN\cite{1703.10593} and GauGAN\cite{1903.07291}. However these models are also significantly more complex and did not seem conducive to a lightweight and performant system with the given requirements. Image generative models are notoriously unstable in training, intuition (and Jonathan Hui) notes that it is easier to recognise an image than it is to create one\cite{hard_to_train_gans}. Therefore traditional NN compression\cite{2006.03669} is often unsuited to GANs. There has been work in GAN compression which has yielded impressive results, but it is a quite esoteric, unconventional compression involving programmatically finding alternate model architectures with neural architecture search (NAS) and teacher--student model pairing to distill knowledge into a smaller architecture\cite{2003.08936}. Simplicity was the first port-of-call for this work thus methods shown in the aforementioned works remain untested in this context for now.

Designing and training a NN is of course not limited to the big name ML frameworks like Tensorflow\cite{tf} or PyTorch\cite{pytorch}, but given the modular nature of a NN and the want in most cases to utilise the GPU for faster training and inference, it is no surprise that a limited number of ML frameworks have become the go-to to create and iterate on a novel ML model and to take care of interfacing with a highly optimized back-end and the GPU at a lower level. Tensorflow offers a relatively full end-to-end pipeline from designing and training a model in Python, to saving and converting a model into something can be run in the browser using TensorflowJS\cite{tfjs}. A trained model is simply a collection of weights and biases and to use this model on a different device or in a different context these weights and biases need to be loaded into a model of the same architecture as was used to train the model. The pipeline from training-to-browser is largely about keeping data structures in order and consistent such that the next process knows how to work with it. As part of the process Tensorflow has developed the TensorflowJS Converter\cite{tfjs_converter} which converts a saved model into a series of binary files which hold the weights and a \verb|.json| file which is the blueprints of the model. When uploading a model TensorflowJS dynamically creates a succession of routines in the chosen backend (WebGL, WebAssembly, WebGPU, etc.) to take care of specific tasks such as 2D convolutions, batchnorm, max pooling etc.

There exists nothing even similar to \textit{Learning to See} in the catalog of examples provided by TensrflowJS --- i.e.\ a model which takes video as input and produces video as output. Thus the uniqueness of this use case has yielded a number of hurdles to overcome to achieve something usable on the web for \textit{most users} (users with a reasonably modern laptop, produced in the past 5 years).

@@

It was found that for a model that takes an input image size of 256$\times$256 pixels and produces an output of the same size, a dataset of 800--1000 images uniformly shuffled with a batch size of 4 was sufficient to produce \textit{interesting} results --- in this case \textit{interesting} means: varied enough to provide unique experiences whilst producing a output which is representative of the input data. Differences between the model sizes can be seen in Tables~\ref{table:numFilters},~\ref{table:model_details} and Figure~\ref{fig:modelSizes}. The models were trained on a NVIDIA Quadro RTX 6000 with 24GB memory locally, or a Tesla P100-PCIE with 16GB memory using Google Colab. In either case a usable model will train in a few hours allowing for a relatively quick development cycle. %More details on the training process and some tools developed to make the process easy and fast in appendix \ref{appendix:training}.

The Pix2Pix generator is an Autoencoder similar to the ``U-Net'' architecture\cite{1505.04597} but differs in some key ways, and we offer some further changes to drastically reduce the size. The generator consists of 8 2D convolution layers downsampling the input data in width and height, and 7 transpose convolutions upsampling the data back to original size. Each layer is batchnormalized\cite{1502.03167} apart from the first, and layers 8--10 have dropout\cite{JMLR:v15:srivastava14a} applied at a rate of 50\%. The Pix2Pix architecture includes no avg- or max-pooling as all filter kernels are 4$\times$4 and a stride of 2 is used. This has the effect of downsampling the layer input by half in width and height, thus circumventing the need for any kind of pooling, otherwise known as an \textit{all convolutional net}\cite{1412.6806}. The difference proposed in this work is to greatly reduce the number of filters used in each layer. The majority of trainable parameters are found in the middle of the generator as a result of the structure described. By reducing the number of filters in each layer by up to a factor of 8, very usable results were obtained whilst reducing the model size --- more discussion on the impacts of a smaller model are found in \textbf{Results}~\ref{section:results}.

@@

Once a model is trained, it is saved using the \verb|HDF5| format (\verb|.h5|) and converted using the TensorflowJS Converter. The Converter provides options for compressing the model of varying levels of severity. It was found that the compression of the model:

@@

Therefore the highest compression of \verb|uint8_affine_quantize| can be used to reduce the size by 4 times, while maintaining only a very slightly compromised output.

@@

Finally, a bespoke WebGL wrapper consisting of a number of shader programs was created to handle the input data (pre-augmentation to match that in training) and the output data to render the image to the screen as video. It was found that the image handling capabilities of TensorflowJS seem to be tailored to static images. In inference, the data is of the \textit{tensor} data structure and is transformed and manipulated by the model, but at each iteration of the draw cycle a frame of video must be preprocessed and then converted into a format which can be digested by the model, and then the output of the model must be converted back into a data structure which the browser can handle to display as an image, or in our case, video. The internal computation at inference is handled by WebGL shader programs compiled by TensorflowJS, but the input/output was a rather slow process of converting the tensor data back to pixels value-by-value.

>>>>

Unfortunately the gallery in itself is a black-box to many, entry restricted by a fee or simply by the location of the gallery. Moving such interactive experiences online is a clear next step as once something is online it is immediately available to anyone with an internet connection, however this then poses numerous challenges. Even a modestly sized NN can contain many millions of parameters. Modern computers are able to perform millions of floating point operations per second (FLOPS) but advances in ML research have shown that \textit{deeper} models and more data produces better results\cite{2005.14165}; as such, without expensive hardware acceleration from discrete graphics processing units (GPU) or even more modern tensor processing units (TPU), an average modern computer cannot crunch the numbers needed to run a large and deep NN\@. Even if one \textit{could} perform the necessary FLOPS, storing this number of parameters can take up gigabytes of memory and so downloading the weights and biases of a pre-trained model can be a time consuming process detracting from the real-time focus.

@@

Getting NNs working online and on the clients machine is not a new thing and with modern CPUs it has certainly been shown possible to train and run a simple enough model in the browser\cite{karparthy_conjs_2016, brainjs, synaptic, mind, kerasjs} and many tools exist which aim to visualise the architecture of a model and show the training process\cite{broad_cnn_vis_2016, tensorflow_playground, 1704.01942}. Such tools are invaluable as learning resources for those dedicated to learning about machine learning. However they are often tailored to that use-case expecting the user to have higher motivation for understanding the process---more than the average online `passer-by'.

@@

\textit{Learning to See} is an adaptation of the Pix2Pix model\cite{1611.07004} which is a Conditional Generative Adversarial Network (cGAN)\cite{1701.00160}. Unlike a conventional GAN\cite{1406.2661}, a cGAN takes in two data samples as input, in the case of the Pix2Pix model it is a \textit{target} image and an \textit{input} image. Pix2Pix also learns a structured loss which means it is able to penalize any possible structure that differs between output and target\cite{1611.07004}. This model architecture is extremely flexible and has been shown to be suitable to many use cases as detailed in the original paper. However neither the original Pix2Pix authors nor \textit{Learning to See} had an incentive to create a small model; a good result is simply a sharp image with a predictable output from a given input. As stated by the Pix2Pix authors measuring the results of an image generating GAN is an open issue\cite{1606.03498, 1611.07004} and in an artistic context quantifying the outcome is largely pointless---a good result is one which delivers the message of the artist.

It is worth noting that other generative models of similar architecture to Pix2Pix have been developed since, notably CycleGAN\cite{1703.10593} and GauGAN\cite{1903.07291}. However these models are also significantly more complex and did not seem conducive to a lightweight and performant system with the given requirements. Image generative models are notoriously unstable in training; intuition (and Jonathan Hui) notes that it is easier to recognise an image than it is to create one\cite{hard_to_train_gans}, therefore traditional NN compression\cite{2006.03669} is often unsuited to GANs. There has been work in GAN compression which has yielded impressive results, but it is a quite esoteric, unconventional compression involving programmatically finding alternate model architectures with neural architecture search (NAS) and teacher--student model pairing to distill knowledge into a smaller architecture\cite{2003.08936}. Simplicity was the first port-of-call for this work thus methods shown in the aforementioned works remain untested in this context for now.

Designing and training a NN is of course not limited to the big name ML frameworks like Tensorflow\cite{tf} or PyTorch\cite{pytorch}, but given the modular nature of a NN and the want in most cases to utilise the GPU for faster training and inference, it is no surprise that a limited number of ML frameworks have become the go-to to create and iterate on a novel ML model and to take care of interfacing with a highly optimized back-end and the GPU at a lower level. Tensorflow offers a relatively full end-to-end pipeline from designing and training a model in Python, to saving and converting a model into something can be run in the browser using TensorflowJS\cite{tfjs}. A trained model is simply a collection of weights and biases and to use this model on a different device or in a different context these weights and biases need to be loaded into a model of the same architecture as was used to train the model. The pipeline from training-to-browser is largely about keeping data structures in order and consistent such that the next process knows how to work with it. As part of the process Tensorflow has developed the TensorflowJS Converter\cite{tfjs_converter} which converts a saved model into a series of binary files which hold the weights and a \verb|JSON| file which is the blueprints of the model. When uploading a model TensorflowJS dynamically creates a succession of routines in the chosen backend (WebGL, WebAssembly, WebGPU, etc.) to take care of specific tasks such as 2D convolutions, batchnorm, max pooling etc.

There exists nothing even similar to \textit{Learning to See} in the catalog of examples provided by TensrflowJS---i.e.\ a model which takes video as input and produces video as output. Thus the uniqueness of this use case has yielded a number of hurdles to overcome to achieve something usable on the web for \textit{most users} (users with a reasonably modern laptop, produced in the past 5 years).

@@

It was found that for a model that takes an input image size of 256$\times$256 pixels and produces an output of the same size, a dataset of 800--1000 images uniformly shuffled with a batch size of 4 was sufficient to produce \textit{interesting} results---in this case \textit{interesting} means: varied enough to provide unique experiences whilst producing an output which is representative of the input data. Differences between the model sizes can be seen in Tables~\ref{table:numFilters},~\ref{table:model_details} and Figure~\ref{fig:modelSizes}. The models were trained on a NVIDIA Quadro RTX 6000 with 24GB memory locally, or a Tesla P100-PCIE with 16GB memory using Google Colab. In either case a usable model will train in a few hours allowing for a relatively quick development cycle.

The Pix2Pix generator is an Autoencoder similar to the ``U-Net'' architecture\cite{1505.04597} but differs in some key ways, and we offer some further changes to drastically reduce the size. The generator consists of 8 2D convolution layers downsampling the input data in width and height, and 7 transpose convolutions upsampling the data back to original size. Each layer is batchnormalized\cite{1502.03167} apart from the first, and layers 8--10 have dropout\cite{JMLR:v15:srivastava14a} applied at a rate of 50\%. The Pix2Pix architecture includes no avg- or max-pooling as all filter kernels are 4$\times$4 and a stride of 2 is used. This has the effect of downsampling the layer input by half in width and height, thus circumventing the need for any kind of pooling, otherwise known as an \textit{all convolutional net}\cite{1412.6806}. The difference proposed in this work is to greatly reduce the number of filters used in each layer. The majority of trainable parameters are found in the middle of the generator as a result of the structure described. By reducing the number of filters in each layer by up to a factor of 8, very usable results were obtained whilst reducing the model size---more discussion on the impacts of a smaller model are found in \textbf{Results}~\ref{section:results}.

@@

Once a model is trained, it is saved using the \verb|HDF5| format (\verb|.h5|) and converted using the TensorflowJS Converter. The converter provides options for compressing the model of varying levels of severity. It was found that the compression of the model:

@@

Therefore the highest compression of \verb|uint8_affine_quantize| can be used to reduce the size by 4 times, while maintaining only a very slightly compromised output. A reduced file size is of great benefit online as it greatly reduces the amount of time to download for the user and the time it takes for TensorflowJS to load the model from the binaries.

@@

Finally, a bespoke WebGL wrapper consisting of a number of shader programs was created to handle the input data (pre-augmentation to match that in training) and the output data to render the image to the screen as video. It was found that the image handling capabilities of TensorflowJS seem to be tailored to static images. In inference the data is of the \textit{tensor} data structure and is transformed and manipulated by the model, but at each iteration of the draw cycle a frame of video must be preprocessed and then converted into a format which can be digested by the model, and then the output of the model must be converted back into a data structure which the browser can handle to display as an image. The internal computation at inference is handled by WebGL shader programs compiled by TensorflowJS, but the input/output was a rather slow process of converting the tensor data back to pixels value-by-value.
